Fusing Filters with Integer Linear Programming

### 2014-05-14 2014betterfusionforfilters/section/04-ILP.tex
add obj function to final, optimised version

\subsubsection{Fusion-preventing path optimisation}
\label{s:OptimisedConstraints}
We can remove certain constraints and variables, by only checking nodes if there is no \emph{fusion-preventing path} between $i$ and $j$.
If there is a fusion-preventing path between two nodes, they are known to be in different clusters.
We can simplify the integer linear program and make the job of the ILP solver easier, by omitting these constraints.

@@

With $split$ defined, we can refine our formulation to only generate constraints between two nodes if they may be fused together.

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j) \not\in E \wedge (j,i) \not\in E$)            \\
\\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j,\fusible) \in E$)     \\
\\
           \>             \>       \> $\pi_i < \pi_j$ \>       \>            \\
           \> (if $split(i,j) \wedge (i,j,\fusionpreventing) \in E$)    \\
\\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (if $split(i,j)$) \\
           \> $c_{i }$    \> $ = $ \> $ 1 $           \>       \>            \\
           \> (if $\neg(split(i,j))$) \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \emptyset$)  \\
           \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \{i',j'\} \wedge$ \\

           \> \> $ split(i,j) \wedge split(i',j) \wedge split(i,j') \wedge split(i',j') $) \\
\end{tabbing}

>>>>

\subsubsection{Fusion-preventing path optimisation}
\label{s:OptimisedConstraints}
We can remove certain constraints and variables, by only checking nodes if there is no \emph{fusion-preventing path} between $i$ and $j$.
If there is a fusion-preventing path between two nodes, they are known to be in different clusters.
With our specific knowledge of the problem domain, we can simplify the integer linear program and make the job of the ILP solver easier by omitting these constraints.
Creating fewer constraints and variables should make it faster for the ILP solver to find a solution.

@@

With $split$ defined, we can refine our formulation to only generate constraints between two nodes if there is a chance they may be fused together.

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \>     \> $\Sigma_{(i,j) \in E} W_{ij} x_{ij}$   \\
           \> (if $split(i,j)$)         \\
           \> $+$ \> $\Sigma_{i \in V} N c_i$  \\
Subject to \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j) \not\in E \wedge (j,i) \not\in E$)            \\
\\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \> (if $split(i,j) \wedge (i,j,\fusible) \in E$)     \\
\\
           \>             \>       \> $\pi_i < \pi_j$ \>       \>            \\
           \> (if $split(i,j) \wedge (i,j,\fusionpreventing) \in E$)    \\
\\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (if $split(i,j)$) \\
           \> $c_{i }$    \> $ = $ \> $ 1 $           \>       \>            \\
           \> (if $\neg(split(i,j))$) \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \emptyset$)  \\
           \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\bot \in \{\iiter_{\Gamma,C}(i), \iiter_{\Gamma,C}(j)\}$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\iiter_{\Gamma,C}(i) \not= \iiter_{\Gamma,C}(j) \wedge parents(i,j) = \{i',j'\} \wedge$ \\

           \> \> $ split(i,j) \wedge split(i',j) \wedge split(i,j') \wedge split(i',j') $) \\
Where      \>                                       \\
           \> $W_{ij} = N^2$ \> $~|$ \> $(i,j) \in E $         \\
           \> \> \> (fusing $i$ and $j$ will reduce memory traffic)         \\
           \> $W_{ij} = N^2$ \> $~|$ \> $\exists k. (k,i) \in E \wedge (k,j) \in E $     \\
           \> \> \> ($i$ and $j$ share an input array)                                         \\
           \> $W_{ij} = 1$   \> $~|$ \> $@otherwise@$                                                  \\
           \> \> \> (the only benefit is loop overhead)                                        \\
           \\
           \> $N = |V|$
\end{tabbing}
